Proof equivalence in \capsabbrev{mll} is \capsabbrev{pspace}-complete

### 2015-07-07 main/LMCS.tex
layout, spacing, "parity problem" comment

Taking a different perspective, we may alternatively consider a switching graph as a directed tree, rooted in the root connective of $\Gamma$. This establishes a bijection between the non-root vertices and the edges, which associates vertex with the edge connecting it to its parent.

@@

We will demonstrate 1), while 2) is similar. Let $\links\perm*\links'$ by rewiring a jump $\jump va$ to $\jump vb$. By fixing a switching for $\Gamma$, we obtain a switching graph for each of the two nets, where the jump is named $v(1)$ in each. There are two possibilities, illustrated below. %in Figure~\ref{fig:rewiring directed switching graph}.
On the left, if the jump $v(1)$ is directed upward, then the target of each edge in the directed switching graphs remains the same---in particular $v(1)$ has target $v$---and the induced permutation is the identity.

\[
	\vc{\parityNetC1} \perm \vc{\parityNetC2}
\qquad\qquad\qquad
	\vc{\parityNetD1} \perm \vc{\parityNetD2} \kern5pt{=}\kern-5pt \vc{\parityNetD3}
\]

%On the right, in Figure~\ref{fig:rewiring directed switching graph},
Above on the right, if the jump from $v$ is directed downward, the subtree of $a$ will get the new root node $b$. Then the vertices that are associated with a new edge are exactly those on the path $\path ab$ in the switching graph, as illustrated below.
%
\[
	\parityNetE
\]
%
Since the connectives in $\Gamma$ were assumed to be strictly alternating, there are an odd number of vertices on this path, $2n+3$: each even $v_i$ must be a $\bot$ or $\pr$, while each odd $v_j$ must be a $\1$ or $\tn$. The permutation induced is then as follows. Since $v$ has target $a$ in the first switching graph, and target $b$ in the second, it takes $a$ to $b$. Further, since an edge connecting $v_i$ and $v_{i+1}$ has target $v_{i+1}$ in the first, but target $v_i$ in the second graph, the permutation takes $v_{i+1}$ to $v_i$. The complete permutation is then a cyclic one taking each vertex on the path $\path ab$ to the previous, and the first to the last. A cyclic permutation of odd length has even parity.

@@

@@

\begin{proposition}
A basic sequent $\1^n,\bigtn(\bot^n)$ is inhabited by exactly $n!$ inequivalent proof nets.
\end{proposition}

>>>>

Taking a different perspective, we may alternatively consider a switching graph as a directed tree, rooted in the root connective of $\Gamma$. This establishes a bijection between the non-root vertices and the edges, which associates each vertex with the edge connecting it to its parent.

%Next, we take a switching graph $G$ for the proof net $(\Gamma,\links)$, and  consider it as a directed tree, rooted in the root connective of $\Gamma$. This establishes a bijection $f_G$ between the non-root vertices and the edges, which associates each vertex with the edge connecting it to its parent: for a given vertex $v$, if $\edge v{\path wx}$ is the path in $G$ from $v$ to the root $x$, then $f_G(v)=a$ is the name of the $\edge vw$ in the above scheme, i.e.\ $a=v(i)$ or $a=w(i)$ for some index $i$.

@@

We will demonstrate 1), while 2) is similar. Let $\links\perm*\links'$ by rewiring a jump $\jump va$ to $\jump vb$. By fixing a switching for $\Gamma$, we obtain a switching graph for each of the two nets, where the jump is named $v(1)$ in each. There are two possibilities, illustrated below. On the left, if the jump $v(1)$ is directed upward, then the target of each edge in the directed switching graphs remains the same---in particular $v(1)$ has target $v$---and the induced permutation is the identity.

\[
	\vc{\parityNetC1} ~\perm~ \vc{\parityNetC2}
\qquad\qquad\qquad
	\vc{\parityNetD1} ~\perm~ \vc{\parityNetD2} ~\kern5pt{=}\kern-5pt~ \vc{\parityNetD3}
\]

Above on the right, if the jump from $v$ is directed downward, the subtree of $a$ will get the new root node $b$. Then the vertices that are associated with a new edge are exactly those on the path $\path ab$ in the switching graph, as illustrated below.
%
\[
	\parityNetE
\]
%
Since the connectives in $\Gamma$ were assumed to be strictly alternating, there are an odd number of vertices on this path, $2n+3$: each even $v_i$ must be a $\bot$ or $\pr$, while each odd $v_j$ must be a $\1$ or $\tn$. The permutation induced is then as follows. Since $v$ has target $a$ in the first switching graph, and target $b$ in the second, it takes $a$ to $b$. Further, since an edge connecting $v_i$ and $v_{i+1}$ has target $v_{i+1}$ in the first, but target $v_i$ in the second graph, the permutation takes $v_{i+1}$ to $v_i$. The complete permutation is then a cyclic one taking each vertex on the path $\path ab$ to the previous, and the first to the last. A cyclic permutation of odd length has even parity.

@@

For our encoding, in Section~\ref{sec:encoding}, this poses the following challenge (the \emph{parity problem}): when we want the encoding to produce proof nets that are equivalent, we must ensure that these do not inadvertently have odd parity.

@@

\begin{proposition}
A basic sequent $\1^n,\bigtn(\bot^n)$ is inhabited by $n!$ inequivalent proof nets.
\end{proposition}
